---
title: "Final project"
author: "Xing Cheng, Jihyun Lee, Hyunjoe Yoo"
date: "11/10/2021"
output: pdf_document
---

[Xing]
```{r}
#install.packages("tidycensus")
library(tidycensus)
library(httr)
library(jsonlite)
#install.packages("sqldf")
library(sqldf)
library(tidyverse)
```

```{r}
# I will figure out API later
res2019 = GET("https://data.medicaid.gov/api/1/metastore/schemas/dataset/items/e36d89c0-f62e-56d5-bc7e-b0adf89262b8?show-reference-ids=false")
res2019=fromJSON(rawToChar(res2019$content))
downloadURL2019=res2019$distribution$data$downloadURL
data2019=read.csv(downloadURL2019)

res2020=GET("https://data.medicaid.gov/api/1/metastore/schemas/dataset/items/fbbe1734-b448-4e5a-bc94-3f8688534741?show-reference-ids=false")
res2020=fromJSON(rawToChar(res2020$content))
downloadURL2020=res2020$distribution$data$downloadURL
data2020=read.csv(downloadURL2020)
```




```{r}
data=sqldf("select * from data2019
      union all
      select * from data2020")
```


```{r}
#Remove dot from column names
names(data) <- gsub("\\.", "", names(data))
names(data)
```
Investigated data using excel, I found the column Source is constant value 
```{r}
SourceAbb=sqldf("
select *, row_number() over (order by FFY,Source) as source_id
from
(select distinct FFY, Source from data) temp")

```

```{r}
MeasureInfo=sqldf("
select distinct MeasureAbbreviation, ReportingProgram, Domain, MeasureName, MeasureType, RateDefinition
from data
              ")
```

```{r}
StateReport=sqldf("select State, MeasureAbbreviation, data.FFY, source_id, Population, Methodology, StateRate, NumberofStatesReporting, Median, BottomQuartile, TopQuartile, Notes, StateSpecificComments, RateUsedinCalculatingStateMeanandMedian
      from data, SourceAbb
      where data.FFY=SourceAbb.FFY and data.Source=SourceAbb.Source")
```


```{r}
#install.packages("usmap")
library(usmap)
library(ggplot2)

State_report_freq_2020=sqldf("
select State as state, count(State) as report_freq
from data
where FFY=2020
group by State
order by report_freq desc
      ")

plot_usmap(data=State_report_freq,values='report_freq',color="red")+
  scale_fill_continuous(low="white",high="blue",name="report freq 2020",label=scales::comma)+
  theme(legend.position='right')
```

```{r}
State_report_freq_2019=sqldf("
select State as state, count(State) as report_freq
from data
where FFY=2019
group by State
order by report_freq desc
      ")

plot_usmap(data=State_report_freq,values='report_freq',color="red")+
  scale_fill_continuous(low="white",high="blue",name="report freq 2019",label=scales::comma)+
  theme(legend.position='right')


```


[Jihyun]
I have two parts, 1: higher is better, made two separate columns 2019 and 2020 to compare. 2: lower is better.
My code will generate new csv files, Hannah will generate the bar graphs to compare two average state rates.

Average state rate for "higher is better" in **2019 and 2020**
```{r fig.show = 'hold', out.width = '50%'}
names(data2019) <- gsub("\\.", "", names(data2019))
#names(data2019)

higher2019=sqldf("select State, avg(StateRate) as StateRate_mean
from data2019
where MeasureType like 'Higher%'
group by State
order by StateRate_mean desc
      ")

par(mar = c(7, 4, 2, 2) + 0.2)
end_point = 0.5 + nrow(higher2019) + nrow(higher2019) - 1
barplot(higher2019$StateRate_mean, col = "grey50",
        main = "",
        ylab = "2019 State Mean (higher is better)", ylim = c(0,70),
        xlab = "",
        space = 1)
text(seq(1.5, end_point, by = 2), par("usr")[3]-0.25,
     srt = 60, adj = 1, xpd = TRUE,
     labels = paste(higher2019$State), cex = 0.65)

names(data2020) <- gsub("\\.", "", names(data2020))

higher2020=sqldf("select State, avg(StateRate) as StateRate_mean
from data2020
where MeasureType like 'Higher%'
group by State
order by StateRate_mean desc
      ")

par(mar = c(7, 4, 2, 2) + 0.2)
end_point = 0.5 + nrow(higher2020) + nrow(higher2020) - 1
barplot(higher2019$StateRate_mean, col = "lightblue",
        main = "",
        ylab = "2020 State Mean (higher is better)", ylim = c(0,70),
        xlab = "",
        space = 1)
text(seq(1.5, end_point, by = 2), par("usr")[3]-0.25,
     srt = 60, adj = 1, xpd = TRUE,
     labels = paste(higher2020$State), cex = 0.65)

h2019 <- higher2019[order(higher2019$State),]
h2020 <- higher2020[order(higher2020$State),]
h2020<- h2020[-40,] #I dropped Pueto Rico because 2019 data doesn't have Pueto Rico.

higher2019_2020<- h2019
higher2019_2020$StateRate_mean_2020<-h2020$StateRate_mean
colnames(higher2019_2020)[2] <- "StateRate_mean_2019"

write.csv(higher2019_2020,"higherbetter_StateRateMean_2019_2020_combine.csv", row.names = FALSE)


```

Average state rate for "lower is better" in 2019 and 2020

```{r fig.show = 'hold', out.width = '32%'}
#names(data2019) <- gsub("\\.", "", names(data2019))
#names(data2019)

lower2019=sqldf("select State, avg(StateRate) as StateRate_mean
from data2019
where MeasureType like 'Lower%'
group by State
order by StateRate_mean desc
      ")

par(mar = c(7, 4, 2, 2) + 0.2)
end_point = 0.5 + nrow(lower2019) + nrow(lower2019) - 1
barplot(lower2019$StateRate_mean, col = "grey50",
        main = "",
        ylab = "2019 State Mean (lower is better)", ylim = c(0,70),
        xlab = "",
        space = 1)
text(seq(1.5, end_point, by = 2), par("usr")[3]-0.25,
     srt = 60, adj = 1, xpd = TRUE,
     labels = paste(lower2019$State), cex = 0.65)

#names(data2020) <- gsub("\\.", "", names(data2020))

lower2020=sqldf("select State, avg(StateRate) as StateRate_mean
from data2020
where MeasureType like 'Lower%'
group by State
order by StateRate_mean desc
      ")

par(mar = c(7, 4, 2, 2) + 0.2)
end_point = 0.5 + nrow(lower2020) + nrow(lower2020) - 1
barplot(lower2019$StateRate_mean, col = "lightblue",
        main = "",
        ylab = "2020 State Mean (lower is better)", ylim = c(0,70),
        xlab = "",
        space = 1)
text(seq(1.5, end_point, by = 2), par("usr")[3]-0.25,
     srt = 60, adj = 1, xpd = TRUE,
     labels = paste(lower2020$State), cex = 0.65)

l2019 <- lower2019[order(lower2019$State),]
l2020 <- lower2020[order(lower2020$State),]
l2020<- l2020[-40,] #I dropped Pueto Rico because 2019 data doesn't have Pueto Rico.

lower2019_2020<- l2019
lower2019_2020$StateRate_mean_2020<-l2020$StateRate_mean
colnames(lower2019_2020)[2] <- "StateRate_mean_2019"

write.csv(lower2019_2020,"lowerbetter_StateRateMean_2019_2020_combine.csv", row.names = FALSE)


```
### Avani
As per a few chosen measures in each domain, how does insurance coverage affect the measures 
reported on per state? What methodology is chosen for the measure, and how does coverage affect 
this? Does it change from 2019 to 2020? 

```{r}
#library(sqldf)
#library(tidyverse)
```

```{r}
data.ins <- sqldf("select * from X2019data
      union
      select * from X2020data")
```
Looking at what the domains are available in the dataset

```{r}
data.ins$Domain <- as.factor(data.ins$Domain)
class(data.ins$Domain)
levels(data.ins$Domain)
```

Looking at what measures are present in domain
```{r}
data.ins$State <- as.factor(data.ins$State)
data.ins$`Reporting Program` <-as.factor(data.ins$`Reporting Program`)
data.ins$`Measure Name` <- as.factor(data.ins$`Measure Name`)
data.ins$`Measure Abbreviation` <- as.factor(data.ins$`Measure Abbreviation`)
data.ins$`Measure Type` <- as.factor(data.ins$`Measure Type`)
data.ins$FFY <- as.factor(data.ins$FFY)
data.ins$Population <- as.factor(data.ins$Population)
data.ins$Methodology <- as.factor(data.ins$Methodology)


```
```{r}
filter(data.ins, Domain == "Behavioral Health Care") %>%
  distinct(`Measure Name`)
```

```{r}
filter(data.ins, Domain == "Dental and Oral Health Services") %>%
  distinct(`Measure Name`)
```
```{r}
filter(data.ins, Domain == "Maternal and Perinatal Health") %>%
  distinct(`Measure Name`)
```
```{r}
filter(data.ins, Domain == "Care of Acute and Chronic Conditions") %>%
  distinct(`Measure Name`)
```
```{r}
filter(data.ins, Domain == "Long-Term Services and Supports") %>%
  distinct(`Measure Name`)
```
```{r}
filter(data.ins, Domain == "Primary Care Access and Preventive Care") %>%
  distinct(`Measure Name`)
```
For the purpose of this project, I'll be looking through a select number of measures. To do this, 
I'll be creating a new table with only these chosen measures

```{r}
chosen.measures <- c("Antidepressant Medication Management: Age 18 and Older", "Follow-Up After Hospitalization for Mental Illness: Age 18 and Older", "Percentage of Eligibles Who Received Preventive Dental Services: Ages 1 to 20", "Contraceptive Care: All Women Ages 21 to 44", "Prenatal and Postpartum Care: Postpartum Care", "Annual Monitoring for Patients on Persistent Medications: Age 18 and Older", "Comprehensive Diabetes Care: Hemoglobin A1c Poor Control (>9.0%): Ages 18 to 75", "National Core Indicators Survey: Age 18 and Older", "Flu Vaccinations for Adults Ages 18 to 64", "Immunizations for Adolescents: Age 13")
data.q <- data.ins %>%
  filter(`Measure Name` %in% chosen.measures)

head(data.q)
```
Now we'll look at the insurance types, which is represented in the "Population" column

```{r}
levels(data.q$Population)
```
CHIP - Childrens Health Insurance Program
Dual Eligibles - dual eligibility for both medicaid and medicare 
For the purposes of this project, we'll look at "Medicaid & CHIP", "Medicaid & CHIP & Dual 
Eligibles", and "Medicaid only"

We notice that not all measures are represented in all states, so we can look into this

```{r}
st.cnt <- data.q %>% 
  filter(`Measure Name` == "Antidepressant Medication Management: Age 18 and Older") %>%
  distinct(State)

st.cntr <- st.cnt$State

length(st.cntr)  
```
In this example, we can see that not all fifty states use this measure. We can look at the rest of 
the measures

```{r}
count.hold <- c()
for (i in 1:10) {
  st.cnt <- data.q %>% 
    filter(`Measure Name` == chosen.measures[i]) %>%
    distinct(State)

  st.cntr <- st.cnt$State

  len.st <- length(st.cntr)
  
  count.hold <- c(count.hold, len.st)
}

st.num <- data.frame(chosen.measures, count.hold)
st.num
```
```{r}
library(ggplot2)
```
```{r}
p1 <- ggplot(data = st.num, aes(x=count.hold, y=chosen.measures)) +
  geom_bar(stat = "identity", color = "light blue", fill = "light blue") +
  geom_text(aes(label=count.hold), vjust=-0.3, size=3.5) +
  labs(y="Measures", x="Count of States Reporting Measure")
  theme_minimal()

p1
ggsave("CountStatesReportingMeasures.png")
```
plot count of measure (out of the ten) by populations
First, we'll look at the different insurance populations 
```{r}
levels(data.q$Population)
```
From here, we'll look at frequency of our measures per populations
```{r}
pops <- c("CHIP only", "Medicaid & CHIP", "Medicaid & Dual Eligibles", "Medicaid", "Medicaid & CHIP & Dual Eligibles", "Medicaid only")

m.hold <- c()

for (i in 1:6) {
  m <- data.q %>% 
    filter(Population == pops[i]) %>%
    distinct(`Measure Name`)
  
  m.cnt <- m$`Measure Name`
  
  len.m <- length(m.cnt)
  
  m.hold <- c(m.hold, len.m)
}

m.num <- data.frame(pops, m.hold)
m.num

p2 <- ggplot(data = m.num, aes(x=m.hold, y=pops)) +
  geom_bar(stat = "identity", color = "light blue", fill = "light blue") +
  geom_text(aes(label=m.hold), vjust=-0.3, size=3.5) +
  labs(y="Governmental Healthcare Populations", x="Count of Marker Measures")
  theme_minimal()
  
p2

ggsave("HealthcarePopulationsPerMeasures.png")
```
Not all coverage populations report on all measures. We can look into this by looking at which 
specific markers are reported bypopulations
```{r}
for (i in 1:6) {
  pop1 <- data.q %>%
    filter(Population == pops[i]) %>%
    distinct(`Measure Name`)
  print(pops[i])
  print(pop1$`Measure Name`)
}
```
We see that Medicaid and CHIP together represent the populations where the most measures were 
taken. We can reason that this is because this is representative of the biggest subset of the
population under government healthcare insurance.


Go into diff types of coverage in governmental insurance 
Within these populations, we see that there are different types of delivery systems of healthcare 
coverage. We can see the effect these types of delivery systems will have on what measures are 
reported by states

```{r}
library(stringr)
library(tidytext)
library(magrittr)
```

```{r}
data.qa <- data.q

data.qa %<>%
  unnest_tokens(word, `State Specific Comments`)
head(data.qa)
```
```{r}
data(stop_words)
data.qa <- data.qa %>%
  anti_join(stop_words)
head(data.qa)
```
```{r}
dev.new(width = 5, height = 10, unit = "cm")
p3 <- data.qa %>%
  count(word, sort = TRUE) %>%
  filter(n > 100) %>%
  filter(n < 900) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word)) +
  geom_col(color = "light blue", fill = "light blue") +
  geom_text(aes(label=n), vjust=-0.3, size=3.5)
  labs(y = NULL) + theme_minimal()

p3

ggsave("CommentWordFrequency.png")
```

State specific comments section is an interesting place to look at what is being said about the 
measures that states report

Words of interest here are "mcos", "ffs", "mco", and "pcc." These are the words that describe the 
different delivery systems. 

by measure and state plot ffs, pccm, etc. -> pick out insurance types 

```{r}
delivery.systems <- c("mcos", "ffs", "mco", "pccm")
```
```{r}

mq1 <- data.q %>% 
    filter(!grepl(delivery.systems == "pccm",`State Specific Comments`)) %>%
    distinct(`Measure Name`)

mql1 <- length(mq1$`Measure Name`)
mql1

mq2 <- data.q %>% 
    filter(!grepl(delivery.systems == "mcos",`State Specific Comments`)) %>%
    distinct(`Measure Name`)

mql2 <- length(mq2$`Measure Name`)
mql2

mq3 <- data.q %>% 
    filter(!grepl(delivery.systems == "mco",`State Specific Comments`)) %>%
    distinct(`Measure Name`)

mql3 <- length(mq3$`Measure Name`)
mql3

mq4 <- data.q %>% 
    filter(!grepl(delivery.systems == "ffs",`State Specific Comments`)) %>%
    distinct(`Measure Name`)

mql4 <- length(mq4$`Measure Name`)
mql4

mshold <- c(mql1, mql2, mql3, mql4)
ms4 <- data.frame(delivery.systems, mshold)

p4 <- ggplot(data = ms4, aes(x=mshold, y=delivery.systems)) +
  geom_bar(stat = "identity", color = "light blue", fill = "light blue") +
  geom_text(aes(label=mshold), vjust=-0.3, size=3.5) +
  labs(y="Delivery System", x="Count of Marker Measures")
  theme_minimal()
  
p4

ggsave("DeliverySystem.png")
```



